DATA 621 Homework #1

### 2019-09-25 hw1/doc/HW1.Rmd

```{r read_data, echo=FALSE}
# Read in the training data
training <- read.csv("../data/moneyball-training-data.csv") %>%
  select(-INDEX) # Dropping meaningless index
# Read in the evaluation data
evaluation <- read.csv("../data/moneyball-evaluation-data.csv")
```

@@

```{r}
quick_summary <- function(df){
  df %>%
    summary() %>%
    kable() %>%
    kable_styling()
}

quick_summary(training)
```

@@

```{r}
remove_bad_values <- function(df){
  # Change 0's to NA so they too can be imputed
  df <- df %>%
  mutate(TEAM_BATTING_SO = ifelse(TEAM_BATTING_SO == 0, NA, TEAM_BATTING_SO))
  # Remove the high pitching strikeout values
  df[which(df$TEAM_PITCHING_SO > 5346),"TEAM_PITCHING_SO"] <- NA
  # Drop the hit by pitcher variable
  df %>%
    select(-TEAM_BATTING_HBP)
}

training <- remove_bad_values(training)
evaluation <- remove_bad_values(evaluation)
```

```{r}
set.seed(42)
knn <- training %>% knnImputation()

impute_values <- function(df, knn){
  impute_me <- is.na(df$TEAM_BATTING_SO)
  df[impute_me,"TEAM_BATTING_SO"] <- knn[impute_me,"TEAM_BATTING_SO"] 
  impute_me <- is.na(df$TEAM_BASERUN_SB)
  df[impute_me,"TEAM_BASERUN_SB"] <- knn[impute_me,"TEAM_BASERUN_SB"] 
  impute_me <- is.na(df$TEAM_BASERUN_CS)
  df[impute_me,"TEAM_BASERUN_CS"] <- knn[impute_me,"TEAM_BASERUN_CS"] 
  impute_me <- is.na(df$TEAM_PITCHING_SO)
  df[impute_me,"TEAM_PITCHING_SO"] <- knn[impute_me,"TEAM_PITCHING_SO"]
  impute_me <- is.na(df$TEAM_FIELDING_DP)
  df[impute_me,"TEAM_FIELDING_DP"] <- knn[impute_me,"TEAM_FIELDING_DP"]
  return(df)
}

training <- impute_values(training, knn)
```

@@

```{r}
add_features <- function(df){
  df %>%
    mutate(TEAM_BATTING_1B = TEAM_BATTING_H - TEAM_BATTING_2B - TEAM_BATTING_3B - TEAM_BATTING_HR)
}

training <- add_features(training)
evaluation <- add_features(evaluation)
```

@@

```{r}
set.seed(42)
train_index <- createDataPartition(training$TARGET_WINS, p = .7, list = FALSE, times = 1)

moneyball_train <- training[train_index,]
moneyball_test <- training[-train_index,]
```

@@

```{r}
full_formula <- "TARGET_WINS ~ TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR + TEAM_BATTING_BB + TEAM_BATTING_SO + TEAM_BASERUN_SB + TEAM_BASERUN_CS + TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_PITCHING_BB + TEAM_PITCHING_SO + TEAM_FIELDING_E + TEAM_FIELDING_DP + TEAM_BATTING_1B + I(TEAM_BATTING_2B^2) + I(TEAM_BATTING_3B^2) + I(TEAM_BATTING_HR^2) + I(TEAM_BATTING_BB^2) + I(TEAM_BATTING_SO^2) + I(TEAM_BASERUN_SB^2) + I(TEAM_BASERUN_CS^2) + I(TEAM_PITCHING_H^2) + I(TEAM_PITCHING_HR^2) + I(TEAM_PITCHING_BB^2) + I(TEAM_PITCHING_SO^2) + I(TEAM_FIELDING_E^2) + I(TEAM_FIELDING_DP^2) + I(TEAM_BATTING_1B^2) + I(TEAM_BATTING_2B^3) + I(TEAM_BATTING_3B^3) + I(TEAM_BATTING_HR^3) + I(TEAM_BATTING_BB^3) + I(TEAM_BATTING_SO^3) + I(TEAM_BASERUN_SB^3) + I(TEAM_BASERUN_CS^3) + I(TEAM_PITCHING_H^3) + I(TEAM_PITCHING_HR^3) + I(TEAM_PITCHING_BB^3) + I(TEAM_PITCHING_SO^3) + I(TEAM_FIELDING_E^3) + I(TEAM_FIELDING_DP^3) + I(TEAM_BATTING_1B^3) + I(TEAM_BATTING_2B^4) + I(TEAM_BATTING_3B^4) + I(TEAM_BATTING_HR^4) + I(TEAM_BATTING_BB^4) + I(TEAM_BATTING_SO^4) + I(TEAM_BASERUN_SB^4) + I(TEAM_BASERUN_CS^4) + I(TEAM_PITCHING_H^4) + I(TEAM_PITCHING_HR^4) + I(TEAM_PITCHING_BB^4) + I(TEAM_PITCHING_SO^4) + I(TEAM_FIELDING_E^4) + I(TEAM_FIELDING_DP^4) + I(TEAM_BATTING_1B^4)"

full_model <- lm(full_formula, moneyball_train)
step_back <- MASS::stepAIC(full_model, direction="backward", trace = F)
poly_call <- summary(step_back)$call
step_back <- lm(poly_call[2], moneyball_train)
summary(step_back)
```

@@

```{r}
moneyball_test$kitchen_sink <- predict(kitchen_sink, moneyball_test)
moneyball_test <- moneyball_test %>%
  mutate(kitchen_sink_error = TARGET_WINS - kitchen_sink)

ggplot(moneyball_test, aes(kitchen_sink_error)) +
  geom_histogram(bins = 50)

summary(moneyball_test$kitchen_sink_error)
```

@@

```{r}
moneyball_test$step_back <- predict(step_back, moneyball_test)
moneyball_test <- moneyball_test %>%
  mutate(step_back_error = TARGET_WINS - step_back)

moneyball_test %>%
  filter(step_back_error > -100) %>%
  ggplot(., aes(step_back_error)) +
  geom_histogram(bins = 50) +
  labs(caption = "Outlier removed")

@@

```{r}
moneyball_test$simple <- predict(simple_fit, moneyball_test)
moneyball_test <- moneyball_test %>%
  mutate(simple_error = TARGET_WINS - simple)

ggplot(moneyball_test, aes(simple_error)) +
  geom_histogram(bins = 50)

summary(moneyball_test$simple_error)
```

>>>>

```{r read_data, echo=FALSE}
# Read in the training data
training <- read.csv("https://raw.githubusercontent.com/crarnouts/Data_621/master/moneyball-training-data.csv") %>%
  select(-INDEX) # Dropping meaningless index
# Read in the evaluation data
evaluation <- read.csv("https://raw.githubusercontent.com/crarnouts/Data_621/master/moneyball-evaluation-data.csv")
```

@@

```{r}
quick_summary <- function(df){
  df %>%
    summary() %>%
    kable() %>%
    kable_styling()
}
quick_summary(training)
```

@@

```{r}
remove_bad_values <- function(df){
  # Change 0's to NA so they too can be imputed
  df <- df %>%
  mutate(TEAM_BATTING_SO = ifelse(TEAM_BATTING_SO == 0, NA, TEAM_BATTING_SO))
  # Remove the high pitching strikeout values
  df[which(df$TEAM_PITCHING_SO > 5346),"TEAM_PITCHING_SO"] <- NA
  # Drop the hit by pitcher variable
  df %>%
    select(-TEAM_BATTING_HBP)
}
training <- remove_bad_values(training)
evaluation <- remove_bad_values(evaluation)
```

```{r}
set.seed(42)
knn <- training %>% knnImputation()
impute_values <- function(df, knn){
  impute_me <- is.na(df$TEAM_BATTING_SO)
  df[impute_me,"TEAM_BATTING_SO"] <- knn[impute_me,"TEAM_BATTING_SO"] 
  impute_me <- is.na(df$TEAM_BASERUN_SB)
  df[impute_me,"TEAM_BASERUN_SB"] <- knn[impute_me,"TEAM_BASERUN_SB"] 
  impute_me <- is.na(df$TEAM_BASERUN_CS)
  df[impute_me,"TEAM_BASERUN_CS"] <- knn[impute_me,"TEAM_BASERUN_CS"] 
  impute_me <- is.na(df$TEAM_PITCHING_SO)
  df[impute_me,"TEAM_PITCHING_SO"] <- knn[impute_me,"TEAM_PITCHING_SO"]
  impute_me <- is.na(df$TEAM_FIELDING_DP)
  df[impute_me,"TEAM_FIELDING_DP"] <- knn[impute_me,"TEAM_FIELDING_DP"]
  return(df)
}
training <- impute_values(training, knn)
```

@@

```{r}
add_features <- function(df){
  df %>%
    mutate(TEAM_BATTING_1B = TEAM_BATTING_H - TEAM_BATTING_2B - TEAM_BATTING_3B - TEAM_BATTING_HR)
}
training <- add_features(training)
evaluation <- add_features(evaluation)
```

@@

```{r}
set.seed(42)
train_index <- createDataPartition(training$TARGET_WINS, p = .7, list = FALSE, times = 1)
moneyball_train <- training[train_index,]
moneyball_test <- training[-train_index,]
```

@@

```{r}
full_formula <- "TARGET_WINS ~ TEAM_BATTING_2B + TEAM_BATTING_3B + TEAM_BATTING_HR + TEAM_BATTING_BB + TEAM_BATTING_SO + TEAM_BASERUN_SB + TEAM_BASERUN_CS + TEAM_PITCHING_H + TEAM_PITCHING_HR + TEAM_PITCHING_BB + TEAM_PITCHING_SO + TEAM_FIELDING_E + TEAM_FIELDING_DP + TEAM_BATTING_1B + I(TEAM_BATTING_2B^2) + I(TEAM_BATTING_3B^2) + I(TEAM_BATTING_HR^2) + I(TEAM_BATTING_BB^2) + I(TEAM_BATTING_SO^2) + I(TEAM_BASERUN_SB^2) + I(TEAM_BASERUN_CS^2) + I(TEAM_PITCHING_H^2) + I(TEAM_PITCHING_HR^2) + I(TEAM_PITCHING_BB^2) + I(TEAM_PITCHING_SO^2) + I(TEAM_FIELDING_E^2) + I(TEAM_FIELDING_DP^2) + I(TEAM_BATTING_1B^2) + I(TEAM_BATTING_2B^3) + I(TEAM_BATTING_3B^3) + I(TEAM_BATTING_HR^3) + I(TEAM_BATTING_BB^3) + I(TEAM_BATTING_SO^3) + I(TEAM_BASERUN_SB^3) + I(TEAM_BASERUN_CS^3) + I(TEAM_PITCHING_H^3) + I(TEAM_PITCHING_HR^3) + I(TEAM_PITCHING_BB^3) + I(TEAM_PITCHING_SO^3) + I(TEAM_FIELDING_E^3) + I(TEAM_FIELDING_DP^3) + I(TEAM_BATTING_1B^3) + I(TEAM_BATTING_2B^4) + I(TEAM_BATTING_3B^4) + I(TEAM_BATTING_HR^4) + I(TEAM_BATTING_BB^4) + I(TEAM_BATTING_SO^4) + I(TEAM_BASERUN_SB^4) + I(TEAM_BASERUN_CS^4) + I(TEAM_PITCHING_H^4) + I(TEAM_PITCHING_HR^4) + I(TEAM_PITCHING_BB^4) + I(TEAM_PITCHING_SO^4) + I(TEAM_FIELDING_E^4) + I(TEAM_FIELDING_DP^4) + I(TEAM_BATTING_1B^4)"
full_model <- lm(full_formula, moneyball_train)
step_back <- MASS::stepAIC(full_model, direction="backward", trace = F)
poly_call <- summary(step_back)$call
step_back <- lm(poly_call[2], moneyball_train)
summary(step_back)
```

@@

```{r}
moneyball_test$kitchen_sink <- predict(kitchen_sink, moneyball_test)
moneyball_test <- moneyball_test %>%
  mutate(kitchen_sink_error = TARGET_WINS - kitchen_sink)
ggplot(moneyball_test, aes(kitchen_sink_error)) +
  geom_histogram(bins = 50)
summary(moneyball_test$kitchen_sink_error)
```

@@

```{r}
moneyball_test$step_back <- predict(step_back, moneyball_test)
moneyball_test <- moneyball_test %>%
  mutate(step_back_error = TARGET_WINS - step_back)
moneyball_test %>%
  filter(step_back >0) %>%
  filter(step_back<182) %>% 
  ggplot(., aes(step_back_error)) +
  geom_histogram(bins = 50) +
  labs(caption = "Outlier removed")

@@

```{r}
moneyball_test$simple <- predict(simple_fit, moneyball_test)
moneyball_test <- moneyball_test %>%
  mutate(simple_error = TARGET_WINS - simple)
ggplot(moneyball_test, aes(simple_error)) +
  geom_histogram(bins = 50)
summary(moneyball_test$simple_error)
```

### Regression Ensemble 
```{r}
moneyball_test <- moneyball_test %>% mutate(ensemble_score = (kitchen_sink + step_back + simple)/3)
moneyball_test <- moneyball_test %>%
  mutate(ensemble_error = TARGET_WINS - ensemble_score)
ggplot(moneyball_test, aes(ensemble_error)) +
  geom_histogram(bins = 50)
summary(moneyball_test$ensemble_error)

```
